---
title: "Covid19"
author: "Rahma Fairuz Rania"
date: "2025-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is project for self learning purpose, reference from 
[Tech Tribe](https://www.youtube.com/watch?v=D_CNmYkGRUc). You can download datasets from [here](https://www.kaggle.com/datasets/sudalairajkumar/novel-corona-virus-2019-dataset/versions/25?resource=download) 

## Libraries

```{r, message=FALSE, warning=FALSE}
library(Hmisc) # for data analysis inc missing value handle, character manipulation
library(dplyr) # data manipulation
library(tidyverse) # data science lib
library(ggplot2) # data visualization
library(glue) # tooltip
library(plotly) # interactive visualization
library(rworldmap) # world map
```



## Import Data

```{r}
rm(list=ls()) # remove all variable stored previously
dt <- read.csv("C:/Users/Asus/Downloads/COVID19_line_list_data.csv")
head(dt)
```

## data wrangling and visualization

```{r}
dim(dt)
```

```{r}
colnames(dt)
```
```{r}
# remove unused column
dt <- dt %>% select(-c( "case_in_country","X","X.1","X.2","X.3","X.4","X.5","X.6","source"))
```


```{r}
colSums(is.na(dt))
```

```{r}
sum(duplicated(dt))
```
Our data has no duplicate data but has missing value. You can also see the data information by describe().


## Data Wrangling and Visualization

```{r}
# death rate
dt$death_count <- as.integer(dt$death != 0) # only take 1 value (died), if its true set it to 1
sum(dt$death_count)/nrow(dt)
```
Death rate is about 0.05%


```{r}
# which age most died
char1 <- dt %>%
  filter(death == 1) %>%
  group_by(age) %>%
  summarise(total = n()) %>%  
  arrange(desc(total)) %>% 
  ungroup() %>% 
  head(5) %>% 
  mutate(label = paste("Age:", age, "\nTotal:", total, "People")) %>%  

  ggplot(aes(x = reorder(as.factor(age), -total), y = total, 
             fill = as.factor(age), text = label)) +  
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Most Common Ages of Death Due to COVID-19",
       x = "Age (Years Old)",
       y = "Total Deaths") +
  theme_classic() +
  theme(legend.position = "none")

ggplotly(char1, tooltip = "text") 
```

Older people can easily infected by corona and most of them were died.

```{r}
# which gender most died
male = subset(dt, gender   == "male")
female = subset(dt, gender == "female")
mean(male$death_count) 
mean(female$death_count)
```
```{r}
by_gender <- rbind(male, female)

char2 <- by_gender %>% group_by(gender) %>% summarise(total = n()) %>% ungroup() %>% mutate(label =paste("Gender:", gender, "\nTotal:", total, "People")) %>% ggplot(aes(x = as.factor(gender), y = total, 
             fill = as.factor(gender), text = label)) +  
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Death by Gender",
       x = "Gender",
       y = "Total Deaths") 

ggplotly(char2, tooltip = "text")
```
Men are more likely died when they affected by corona virus.


```{r,  warning=FALSE}
# total cases per country
case_country <- dt %>% group_by(country) %>% summarise(total = n()) %>% arrange(desc(total))
case_country

# Join data with map
map_data <- joinCountryData2Map(case_country, joinCode = "NAME", nameJoinColumn = "country")

# Plot the map
mapCountryData(map_data, nameColumnToPlot = "total", 
               mapTitle = "COVID-19 Cases by Country", 
               colourPalette = "heat", catMethod = "fixedWidth")

```

The most infected in the early 2020 is in East Asian such China, Japan, and Soouth Korea.